\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, shapes.geometric, fit, calc, decorations.pathreplacing, patterns, shadows.blur}

\geometry{margin=1in}

\title{LSH-SFA: Spacetime Field Attention with Wave Packets, Light-Cone Causality, and PDE-Driven Evolution}

\author{Hengbo Li\\
\texttt{1147502779@qq.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
We present LSH-SFA (Spacetime Field Attention), a physics-informed neural architecture that replaces Transformer's core mechanisms with physically grounded alternatives. At the attention level, LSH-SFA replaces QKV dot-product with wave packet representations $(k, w, \omega, \theta)$, achieving $O(n)$ linear complexity via Gaussian similarity with learnable bandwidth. At the causality level, hard causal masking is replaced by light-cone attention with frequency-adaptive exponential decay, where low-frequency tokens propagate globally and high-frequency tokens decay locally. At the feed-forward level, LSH-SFA replaces FFN with PDE-driven field evolution: diffusion ($\partial F/\partial t = \alpha \nabla^2 F + \beta F^3$) for classification and convection ($\partial F/\partial t = -c \partial F/\partial x + \beta F^3$) for generation, inspired by the Ginzburg-Landau equation. A resonance coupling mechanism modulates field interactions through frequency-dependent gating. We provide a stability analysis showing that the discrete PDE scheme is stable when $\alpha \leq h^2/(2d)$ and $|\beta| \leq 1/M^2$ with field clamping to $[-M, M]$. On AG News text classification, a 2-layer LSH-SFA with 1.02M parameters achieves 91.80\% accuracy, matching a 4-layer Transformer while using 14\% fewer parameters. On WikiText-2 language modeling, LSH-SFA achieves perplexity of 654 vs. the Transformer's 1069 (39\% improvement), with no overfitting observed. The neural architecture serves as the computational foundation for the LSH semantic framework~\cite{lsh-three-layer}.
\end{abstract}

\section{Introduction}

\subsection{The Limitations of Transformer Architecture}

The Transformer architecture \citep{vaswani2017attention} has three fundamental limitations rooted in its design choices:

\textbf{Quadratic Attention.} QKV dot-product attention has $O(n^2)$ complexity, limiting sequence length. Linear attention methods \citep{katharopoulos2020transformers, choromanski2021rethinking} reduce complexity but sacrifice expressiveness through kernel approximations.

\textbf{Hard Causal Masking.} The binary mask treats all past tokens equally---it cannot distinguish ``near past'' from ``distant past.'' This contradicts the physical principle that information propagates with distance-dependent attenuation, as characterized by light cones in spacetime physics.

\textbf{Uninterpreted Feed-Forward.} The FFN has no physical interpretation and requires $4 \times d_{\text{model}}$ hidden dimensions, contributing the majority of parameters without providing task-appropriate inductive biases.

\subsection{Our Approach: Physics-Informed Neural Architecture}

LSH-SFA addresses these limitations by replacing each Transformer component with a physically grounded alternative:

\begin{enumerate}
\item \textbf{Wave Packets replace QKV:} Each token is represented as a wave packet $(k, w, \omega, \theta)$, where $k$ is the wave vector (semantic position), $w$ is the weight (content), $\omega$ is the angular frequency (temporal dynamics), and $\theta$ is the phase (context alignment)
\item \textbf{Light-Cone Attention replaces hard masking:} Frequency-adaptive exponential decay provides continuous, physically motivated causality
\item \textbf{PDE Evolution replaces FFN:} Diffusion for classification, convection for generation, with energy normalization for stability
\end{enumerate}

The neural architecture presented here serves as the computational foundation for the LSH semantic framework~\cite{lsh-three-layer}, which provides the element-observers-world architecture for knowledge representation and perception.

\subsection{Contributions}

Our main contributions are:

\begin{enumerate}
\item We propose wave packet representations $(k, w, \omega, \theta)$ that provide physically interpretable token embeddings with explicit temporal dynamics
\item We introduce light-cone attention with frequency-adaptive decay, achieving $O(n)$ causal complexity while preserving long-range information flow for low-frequency tokens
\item We replace FFN with PDE-driven field evolution inspired by the Ginzburg-Landau equation, with separate diffusion and convection modes for classification and generation, and provide a stability analysis for the discrete scheme
\item We present resonance coupling for frequency-dependent information transfer between tokens
\item We demonstrate a unified decoder-as-encoder architecture where a single model serves classification, regression, and generation through per-layer causal toggling and PDE mode selection, validated on AG News (91.80\% accuracy with 14\% fewer parameters) and WikiText-2 (39\% perplexity improvement with no overfitting)
\end{enumerate}

\section{Wave Packet Representation}

Each token is represented as a wave packet $(k, w, \omega, \theta)$:

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    wave node/.style={circle, draw, minimum size=1cm, font=\tiny},
    arrow/.style={-{Stealth[length=2mm]}, thick}
]

\node[wave node, fill=blue!20] (k) at (0,0) {$k$};
\node[wave node, fill=green!20] (w) at (2,0) {$w$};
\node[wave node, fill=orange!20] (omega) at (4,0) {$\omega$};
\node[wave node, fill=red!20] (theta) at (6,0) {$\theta$};

\node[below=0.2cm of k, font=\tiny] {Wave Vector\\(Semantic Position)};
\node[below=0.2cm of w, font=\tiny] {Weight\\(Content)};
\node[below=0.2cm of omega, font=\tiny] {Angular Freq.\\(Temporal Dynamics)};
\node[below=0.2cm of theta, font=\tiny] {Phase\\(Context Alignment)};

\draw[arrow] (k) -- (w);
\draw[arrow] (w) -- (omega);
\draw[arrow] (omega) -- (theta);

\end{tikzpicture}
\caption{Wave packet parameters and their physical meanings.}
\label{fig:wave-packet}
\end{figure}

The WavePacketProjector projects hidden states $\mathbf{x}_i \in \mathbb{R}^{d_{\text{field}}}$ to wave packets:

\begin{equation}
[k_i, w_i, \omega_i, \theta_i] = \text{Linear}(\mathbf{x}_i)
\end{equation}

In causal mode, direction bias is applied to $k$:

\begin{equation}
k_i^{\text{causal}} = k_i + b_{\text{dir}} \cdot \frac{i}{\text{seq\_len}}
\end{equation}

This ensures future tokens have larger $k$ values, naturally attenuating via Gaussian similarity.

\section{Field Attention: Light-Cone Mechanism}

\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.6,
    cell/.style={minimum size=0.7cm, draw, font=\tiny},
    decaycell/.style={minimum size=0.7cm, draw, font=\tiny},
]

\node[font=\small\bfseries] at (3.5, 6.5) {Transformer: Hard Mask};
\node[font=\small\bfseries] at (11.5, 6.5) {LSH-SFA: Light Cone};

\foreach \i in {0,...,4} {
    \foreach \j in {0,...,4} {
        \pgfmathtruncatemacro{\val}{\i >= \j ? 1 : 0}
        \ifnum\val=1
            \node[cell, fill=blue!30] at (\j*1.4, 5.0-\i*1.0) {1};
        \else
            \node[cell, fill=red!10] at (\j*1.4, 5.0-\i*1.0) {0};
        \fi
    }
    \node[font=\tiny, left] at (-0.5, 5.0-\i*1.0) {$t_{\pgfmathprintnumber{\i}}$};
}

\foreach \i in {0,...,4} {
    \foreach \j in {0,...,4} {
        \pgfmathtruncatemacro{\val}{\i >= \j ? 1 : 0}
        \ifnum\val=1
            \pgfmathsetmacro{\dist}{\i-\j}
            \pgfmathsetmacro{\decay}{max(0.05, 1.0 - \dist*0.22)}
            \pgfmathtruncatemacro{\intensity}{\decay*100}
            \node[decaycell, fill=blue!\intensity!white] at (8+\j*1.4, 5.0-\i*1.0) {\pgfmathprintnumber[fixed, precision=2]{\decay}};
        \else
            \node[decaycell, fill=gray!5] at (8+\j*1.4, 5.0-\i*1.0) {};
        \fi
    }
    \node[font=\tiny, left] at (7.5, 5.0-\i*1.0) {$t_{\pgfmathprintnumber{\i}}$};
}

\node[font=\tiny, text=red!60!black, align=center] at (3.5, -0.5) {Binary 0/1 mask};
\node[font=\tiny, text=blue!60!black, align=center] at (11.5, -0.5) {Continuous decay};

\draw[-{Stealth}, thick, green!60!black] (6.2, 2.5) -- (7.3, 2.5) node[midway, above, font=\tiny] {replace};

\end{tikzpicture}
\caption{Comparison: Transformer's hard mask vs. LSH-SFA's light-cone attention.}
\label{fig:lightcone}
\end{figure}

LSH-SFA replaces hard masking with frequency-adaptive exponential decay:

\begin{equation}
\gamma_t = \exp\!\left(-\left(\gamma_{\min} + (\gamma_{\max} - \gamma_{\min}) \cdot \sqrt{\frac{|\omega_t|}{|\omega_t| + 1}}\right)\right)
\end{equation}

Low-frequency tokens propagate globally; high-frequency tokens decay locally. This provides spacetime consistency: information propagation respects the causal structure of the sequence, with distance-dependent attenuation governed by the token's temporal frequency.

The cumulative attention is computed via parallel prefix scan in $O(\log n)$ parallel steps:

\begin{equation}
\text{cumKV}_t = \gamma_t \cdot \text{cumKV}_{t-1} + \phi(k_t) \cdot w_t
\end{equation}

\section{PDE-Driven Field Evolution}

LSH-SFA replaces FFN with PDE-driven field evolution. The nonlinear term $\beta F^3$ is inspired by the Ginzburg-Landau equation~\citep{ginzburg1950k}, which models phase transitions in condensed matter physics and provides principled nonlinear feature interaction.

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    node distance=0.6cm and 1.0cm,
    pdebox/.style={rectangle, draw, rounded corners=5pt, minimum height=2.8cm, minimum width=3.2cm, align=center, font=\small, blur shadow},
    fieldnode/.style={circle, draw, fill=#1, minimum size=0.45cm, inner sep=0pt},
    arr/.style={-{Stealth[length=2mm]}, thick, #1},
    biarr/.style={{Stealth[length=2mm]}-{Stealth[length=2mm]}, thick, #1},
    label/.style={font=\scriptsize, align=center},
    eqlabel/.style={font=\scriptsize, align=center, text=#1},
]

\node[pdebox, fill=blue!8] (diff) at (0, 0) {};
\node[label, above=0.1cm] at (diff.north) {\textbf{Diffusion}};
\node[label] at (diff.center) {Classification};

\foreach \i in {0,...,4} {
    \node[fieldnode=blue!30] (d\i) at (-1.0+\i*0.5, -0.4) {};
}
\draw[biarr=blue!60] (d0) -- (d1);
\draw[biarr=blue!60] (d1) -- (d2);
\draw[biarr=blue!60] (d2) -- (d3);
\draw[biarr=blue!60] (d3) -- (d4);
\node[eqlabel=blue!70!black, below=0.15cm] at (diff.south) {$\partial_t F = \alpha\nabla^2 F + \beta F^3$};

\node[pdebox, fill=red!8] (conv) at (5, 0) {};
\node[label, above=0.1cm] at (conv.north) {\textbf{Convection}};
\node[label] at (conv.center) {Generation};

\foreach \i in {0,...,4} {
    \node[fieldnode=red!30] (c\i) at (4.0+\i*0.5, -0.4) {};
}
\draw[arr=red!60] (c0) -- (c1);
\draw[arr=red!60] (c1) -- (c2);
\draw[arr=red!60] (c2) -- (c3);
\draw[arr=red!60] (c3) -- (c4);
\node[eqlabel=red!70!black, below=0.15cm] at (conv.south) {$\partial_t F = -c\,\partial_x F + \beta F^3$};

\node[pdebox, fill=green!8] (wave) at (10, 0) {};
\node[label, above=0.1cm] at (wave.north) {\textbf{Wave}};
\node[label] at (wave.center) {Oscillation};

\foreach \i in {0,...,4} {
    \node[fieldnode=green!30] (w\i) at (9.0+\i*0.5, -0.4) {};
}
\draw[biarr=green!60] (w0) -- (w1);
\draw[biarr=green!60] (w1) -- (w2);
\draw[biarr=green!60] (w2) -- (w3);
\draw[biarr=green!60] (w3) -- (w4);
\node[eqlabel=green!70!black, below=0.15cm] at (wave.south) {$\partial_t^2 F = c^2 \nabla^2 F$};

\node[rectangle, draw, rounded corners=3pt, fill=orange!10, minimum height=1.0cm, minimum width=3.0cm, align=center, font=\small, blur shadow] (res) at (5, -3.2) {\textbf{Resonance Coupling}\\$\text{mod} = e^{-\frac{(\bar{\theta}\cdot\tanh\bar{\omega})^2}{2}} \cdot \tanh\bar{\omega} \cdot \sigma_r$};

\node[rectangle, draw, rounded corners=3pt, fill=purple!10, minimum height=0.8cm, minimum width=3.0cm, align=center, font=\small, blur shadow] (norm) at (5, -5.0) {\textbf{Energy Normalization}\\$F \leftarrow F \cdot \sqrt{E_{\text{target}} / (\text{RMS}(F^2)+\epsilon)}$};

\draw[arr=blue!50] (diff.south) -- (res.north west);
\draw[arr=red!50] (conv.south) -- (res.north);
\draw[arr=green!50] (wave.south) -- (res.north east);
\draw[arr=purple!50] (res.south) -- (norm.north);

\draw[decorate, decoration={brace, amplitude=6pt, mirror}] (-2.0, 1.8) -- (12.0, 1.8) node[midway, above=8pt, font=\small\bfseries] {Spacetime Engine: Three PDE Modes};

\end{tikzpicture}
\caption{Overview of the Spacetime Engine. Three PDE modes provide physically grounded field evolution for different task types. Resonance coupling modulates inter-token information transfer based on frequency similarity, while energy normalization ensures training stability.}
\label{fig:engine-overview}
\end{figure*}

\subsection{Diffusion Equation (Classification)}

For classification tasks, the field evolves according to:

\begin{equation}
\frac{\partial F}{\partial t} = \alpha \nabla^2 F + \beta F^3
\end{equation}

The Laplacian is discretized via centered differences:

\begin{equation}
\nabla^2 F[i] = F[i-1] + F[i+1] - 2F[i]
\end{equation}

\textbf{Physical Interpretation.} The diffusion equation models heat conduction with nonlinear self-interaction:
\begin{itemize}
\item $\alpha \nabla^2 F$: Spatial smoothing (bidirectional aggregation)
\item $\beta F^3$: Nonlinear self-interaction, inspired by the Ginzburg-Landau equation~\citep{ginzburg1950k}
\end{itemize}

\textbf{Connection to Ginzburg-Landau Theory.} The cubic term $\beta F^3$ originates from the Ginzburg-Landau free energy expansion~\citep{ginzburg1950k}, which describes phase transitions in condensed matter physics. In the original framework, the free energy functional $\mathcal{F}[F] = \int \left[\alpha |\nabla F|^2 + \frac{\beta}{2} F^4\right] dx$ yields the gradient flow $\partial F/\partial t = -\delta\mathcal{F}/\delta F = \alpha \nabla^2 F + \beta F^3$. In LSH-SFA, this term provides nonlinear feature interaction that complements the linear Laplacian smoothing, enabling the field to develop non-trivial patterns rather than converging to uniform states. Unlike the standard Ginzburg-Landau equation which includes a linear term $-\gamma F$ for stability, LSH-SFA relies on energy normalization (Section~6) to bound the field magnitude.

\subsection{Convection Equation (Generation)}

For generation tasks, the field evolves according to:

\begin{equation}
\frac{\partial F}{\partial t} = -c \frac{\partial F}{\partial x} + \beta F^3
\end{equation}

The spatial derivative uses backward differences:

\begin{equation}
\frac{\partial F}{\partial x}[i] = F[i] - F[i-1]
\end{equation}

\textbf{Physical Interpretation.} The convection equation models fluid flow with nonlinear self-interaction:
\begin{itemize}
\item $-c \cdot \partial F/\partial x$: Left-to-right advection (information flows from past to future)
\item $\beta F^3$: Nonlinear self-interaction (same Ginzburg-Landau origin as diffusion)
\item Natural causal structure: backward differences ensure no information flows from future to past
\end{itemize}

\textbf{Causal Stability.} The backward difference $\partial F/\partial x[i] = F[i] - F[i-1]$ is an upwind scheme for positive convection speed $c > 0$, which is unconditionally stable for the linear part. Combined with the same $\beta$ stability bound as diffusion ($|\beta| \leq 1/M^2$ with $M = 2$), the convection step preserves causality while allowing nonlinear feature development.

\subsection{Wave Equation (Oscillation)}

For oscillatory dynamics:

\begin{equation}
\frac{\partial^2 F}{\partial t^2} = c^2 \nabla^2 F
\end{equation}

The wave equation models oscillatory phenomena where information propagates as waves rather than diffusing or advecting. Potential applications include periodic patterns (seasonal variations, musical structures), resonance detection (complementing the resonance coupling mechanism), and wave interference in semantic space.

\textbf{Current status.} The wave equation is defined in the Spacetime Engine but not yet used in the current experiments. Its integration is left as future work, requiring careful design of the second-order time discretization and boundary conditions. A natural approach would be the leapfrog scheme $F^{t+1} = 2F^t - F^{t-1} + c^2 \nabla^2_h F^t$.

\subsection{Stability Analysis}

The stability of the discrete PDE steps depends on the relationship between $\alpha$, $\beta$, and the field magnitude:

\begin{proposition}
For the discrete diffusion step $F^{t+1} = F^t + \alpha \nabla^2_h F^t + \beta (F^t)^3$ with field values clamped to $[-M, M]$, the scheme is stable when:
\begin{equation}
\alpha \leq \frac{h^2}{2d}, \quad |\beta| \leq \frac{1}{M^2}
\end{equation}
where $h$ is the grid spacing and $d$ is the spatial dimension.
\end{proposition}

\begin{proof}[Proof sketch]
The Laplacian discretization has spectral radius $\rho(\nabla^2_h) = 4d/h^2$ in $d$ dimensions. For the linear part, the CFL condition requires $\alpha \rho \leq 2$, giving $\alpha \leq h^2/(2d)$. For the nonlinear part, the Jacobian of $\beta F^3$ is $3\beta F^2$, which is bounded by $3|\beta| M^2$. Requiring $3|\beta| M^2 \leq 1$ for contractivity gives $|\beta| \leq 1/(3M^2)$. In practice, we use the more conservative bound $|\beta| \leq 1/M^2$ and clamp $M = 2$, yielding $|\beta| \leq 0.25$. The default initialization $\beta = 0.01$ satisfies this condition with significant margin.
\end{proof}

In the code implementation, the field is clamped to $[-2, 2]$ before computing $F^3$, and energy normalization provides an additional stability guarantee by rescaling the field after each step.

\subsection{Multi-Dimensional Spatial Fields}

A fundamental advantage of the PDE-driven architecture is that the Laplacian operator $\nabla^2$ naturally generalizes across spatial dimensions, enabling LSH-SFA to process data in 1D, 2D, and 3D without architectural changes. This contrasts with Transformers, which are inherently 1D sequential models and require ad-hoc adaptations (e.g., ViT's patch tokenization) for spatial data.

\textbf{Dimensional Generalization of the Laplacian.} The discrete Laplacian in $d$ dimensions is:

\begin{equation}
\nabla^2_h F[\mathbf{i}] = \sum_{j=1}^{d} \left(F[\mathbf{i} + \mathbf{e}_j] + F[\mathbf{i} - \mathbf{e}_j] - 2F[\mathbf{i}]\right)
\end{equation}

where $\mathbf{e}_j$ is the unit vector along dimension $j$. The concrete forms are:

\begin{itemize}
\item \textbf{1D (Text):} $\nabla^2 F[i] = F[i-1] + F[i+1] - 2F[i]$, bidirectional token aggregation
\item \textbf{2D (Image):} $\nabla^2 F[i,j] = F[i\!-\!1,j] + F[i\!+\!1,j] + F[i,j\!-\!1] + F[i,j\!+\!1] - 4F[i,j]$, spatial smoothing across image patches
\item \textbf{3D (Volume/Video):} $\nabla^2 F[i,j,k]$ with 6 neighbors, volumetric diffusion or spatiotemporal propagation
\end{itemize}

The stability condition $\alpha \leq h^2/(2d)$ naturally adapts: higher-dimensional fields require smaller diffusion coefficients, which is physically consistent---more neighbors mean faster information spread.

\textbf{Wave Packet Representation Across Dimensions.} The wave vector $k$ generalizes from a scalar direction bias (1D) to a spatial frequency vector:

\begin{table}[h]
\centering
\caption{Wave packet parameters across spatial dimensions.}
\label{tab:wave_packet_dims}
\begin{tabular}{lcccc}
\toprule
\textbf{Dimension} & $k$ (wave vector) & $\omega$ (frequency) & $\theta$ (phase) & \textbf{Application} \\
\midrule
1D & Scalar direction & Temporal & Context & Text sequences \\
2D & 2D spatial freq. & Temporal & Context & Images, surfaces \\
3D & 3D spatial freq. & Spatiotemporal & Context & Volumes, video \\
\bottomrule
\end{tabular}
\end{table}

In 2D, $k = (k_x, k_y)$ encodes spatial frequency, allowing the Gaussian similarity $A_{ij} = \exp(-\|k_i - k_j\|^2 / 2\sigma^2)$ to capture both horizontal and vertical relationships between image patches. In 3D, $k = (k_x, k_y, k_z)$ extends this to volumetric neighborhoods.

\textbf{Comparison with Transformer Spatial Adaptations.}

\begin{table}[h]
\centering
\caption{Spatial processing: Transformer vs.\ LSH-SFA.}
\label{tab:spatial_comparison}
\begin{tabular}{lcc}
\toprule
\textbf{Property} & \textbf{Transformer} & \textbf{LSH-SFA} \\
\midrule
Spatial encoding & Appended (sin/cos) & Intrinsic (wave vector $k$) \\
1D $\to$ 2D adaptation & Patch tokenization & Native 2D Laplacian \\
1D $\to$ 3D adaptation & No standard method & Native 3D Laplacian \\
Spatial inductive bias & None (learned from data) & Diffusion smoothness \\
Neighbor structure & Global attention & Local $+$ global (PDE) \\
Dimension scaling & $O(n^2)$ regardless & $O(n \cdot 2d)$ neighbors \\
\bottomrule
\end{tabular}
\end{table}

Transformers process images by flattening 2D patches into a 1D sequence, discarding spatial adjacency---two patches that are spatially adjacent may become distant in the token sequence. LSH-SFA preserves spatial structure natively through the Laplacian, which operates on the actual grid topology. This provides a strong inductive bias for spatial coherence: the diffusion PDE naturally enforces smoothness between spatially adjacent elements, while the cubic term $\beta F^3$ allows for sharp feature boundaries.

\textbf{Current Status.} The current implementation operates in 1D for text tasks. Extending to 2D/3D requires: (1) replacing the 1D centered difference with the $d$-dimensional discrete Laplacian, (2) extending $k$ from a scalar to a $d$-dimensional vector in the wave packet projector, and (3) implementing 2D/3D positional encoding for image patches or volumetric grids. The mathematical framework and stability guarantees are already in place.

\section{Resonance Coupling}

Resonance coupling enhances information transfer between tokens with similar oscillation patterns:

\begin{equation}
\text{gate} = \tanh(\bar{\omega})
\end{equation}
\begin{equation}
\text{resonance} = \exp\!\left(-\frac{(\bar{\theta} \cdot \text{gate})^2}{2}\right)
\end{equation}
\begin{equation}
\text{modulation} = \text{resonance} \cdot \text{gate} \cdot \sigma_r
\end{equation}
\begin{equation}
\text{output} = \text{Linear}(F \cdot (1 + \text{modulation}))
\end{equation}

where $\bar{\omega}$ and $\bar{\theta}$ are the mean angular frequency and phase across the time dimension.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=0.8cm and 1.5cm,
    wp/.style={rectangle, draw, rounded corners=3pt, minimum height=1.0cm, minimum width=1.8cm, align=center, font=\scriptsize},
    op/.style={rectangle, draw, rounded corners=3pt, fill=orange!15, minimum height=0.8cm, minimum width=1.5cm, align=center, font=\scriptsize},
    arr/.style={-{Stealth[length=2mm]}, thick}
]

\node[wp, fill=blue!15] (f1) at (0, 1.0) {$\omega_1 = 0.3$\\low freq};
\node[wp, fill=blue!25] (f2) at (0, -1.0) {$\omega_2 = 0.5$\\low freq};
\node[wp, fill=red!15] (f3) at (0, -3.0) {$\omega_3 = 5.0$\\high freq};

\node[op] (gate) at (3, 0) {$\tanh(\bar{\omega})$\\gate};
\node[op] (res) at (5.5, 0) {$e^{-\frac{(\bar{\theta}\cdot g)^2}{2}}$\\resonance};
\node[op, fill=green!15] (mod) at (8, 0) {$F \cdot (1+\text{mod})$\\output};

\draw[arr] (f1.east) -- (gate.west);
\draw[arr] (f2.east) -- (gate.west);
\draw[arr, dashed, gray] (f3.east) -- + (1.5, 0) node[midway, above, font=\tiny, text=gray] {weak};
\draw[arr] (gate) -- (res);
\draw[arr] (res) -- (mod);

\node[font=\tiny, text=blue!60!black, below=0.3cm of f2] {Similar $\omega$ $\rightarrow$ Strong coupling};
\node[font=\tiny, text=red!60!black, below=0.3cm of f3] {Dissimilar $\omega$ $\rightarrow$ Weak coupling};

\end{tikzpicture}
\caption{Resonance coupling mechanism: tokens with similar angular frequencies produce stronger coupling through frequency-dependent gating, while dissimilar frequencies are attenuated.}
\label{fig:resonance}
\end{figure}

\section{Energy Normalization}

Energy normalization prevents gradient explosion:

\begin{equation}
F \leftarrow F \cdot \sqrt{\frac{E_{\text{target}}}{\text{RMS}(F^2) + \epsilon}}
\end{equation}

The scale is clamped to $[0.2, 5.0]$ for stability. In causal mode, normalization is applied per-position; in bidirectional mode, globally. This mechanism serves a dual role: it prevents the PDE evolution from producing unbounded field values, and it acts as an implicit regularizer that eliminates overfitting (as demonstrated in our experiments).

\section{Connection to Semantic Architecture}

The LSH-SFA neural architecture serves as the computational foundation for the LSH semantic framework~\cite{lsh-three-layer}. In that framework, everything is an Element with objective properties; Observers define perception rules that compute subjective properties; and CMD\_PERCEIVE executes perception computations. The wave packet parameters $(k, w, \omega, \theta)$ provide a natural neural realization of the Element model: $k$ encodes semantic position, $w$ encodes content, $\omega$ encodes temporal dynamics, and $\theta$ encodes context alignment. The emergent scheduling mechanism in the semantic framework---where rule-based scoring replaces commanded scheduling---complements the frequency-adaptive decay in light-cone attention: both derive global behavior from local rules rather than centralized control. A detailed presentation of the semantic architecture is provided in~\cite{lsh-three-layer}.

\section{Architecture Overview}

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    node distance=0.6cm and 1.2cm,
    block/.style={rectangle, draw, rounded corners=3pt, minimum height=1.0cm, minimum width=2.4cm, align=center, font=\small},
    smallblock/.style={rectangle, draw, rounded corners=2pt, minimum height=0.7cm, minimum width=1.8cm, align=center, font=\scriptsize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
    dasharrow/.style={-{Stealth[length=2mm]}, dashed, gray},
    label/.style={font=\scriptsize, text=gray},
]

\node[block, fill=blue!10] (input) {Raw Bytes\\$b_i \in \{0..255\}$};

\node[block, fill=blue!20, right=of input] (struct) {Structural\\Encoding};
\node[label, above=0.1cm of struct] {$s \cdot \text{Embed}(b_i) + \text{PosEnc}(i)$};

\node[block, fill=orange!15, right=of struct] (wp) {WavePacket\\Projector};
\node[label, above=0.1cm of wp] {$[k, w, \omega, \theta] = \text{Linear}(\mathbf{x})$};

\node[block, fill=red!15, right=1.5cm of wp] (attn) {Field\\Attention};
\node[label, above=0.1cm of attn] {Gaussian similarity};

\node[block, fill=green!15, right=of attn] (res) {Resonance\\Coupling};
\node[label, above=0.1cm of res] {$F \cdot (1 + \text{mod})$};

\node[block, fill=yellow!20, right=of res] (evo) {Field\\Evolution};
\node[label, above=0.1cm of evo] {PDE: $\alpha\nabla^2 F + \beta F^3$};

\draw[arrow] (input) -- (struct);
\draw[arrow] (struct) -- (wp);
\draw[arrow] (wp) -- (attn) node[midway, above, font=\scriptsize] {$k, w, \omega$};
\draw[arrow] (attn) -- (res);
\draw[arrow] (res) -- (evo);

\node[smallblock, fill=gray!10, below=0.5cm of attn] (norm1) {LayerNorm};
\node[smallblock, fill=gray!10, below=0.5cm of res] (norm2) {LayerNorm};
\node[smallblock, fill=gray!10, below=0.5cm of evo] (norm3) {LayerNorm};

\draw[dasharrow] (attn) -- (norm1);
\draw[dasharrow] (res) -- (norm2);
\draw[dasharrow] (evo) -- (norm3);

\node[draw, dashed, rounded corners=5pt, fit=(wp)(attn)(res)(evo)(norm1)(norm2)(norm3), inner sep=0.3cm, label={[font=\small]above:SpacetimeFieldLayer $\times N$}] (layer) {};

\node[block, fill=purple!15, right=1.5cm of evo] (decoder) {WavePacket\\Decoder};
\draw[arrow] (evo) -- (decoder);

\node[block, fill=blue!10, right=of decoder] (output) {Output\\(logits)};

\draw[arrow] (decoder) -- (output);

\node[font=\scriptsize, text=red!60!black] at ($(struct)!0.5!(wp) + (0,-1.3)$) {Residual $+$ LayerNorm};

\node[draw, dashed, rounded corners=3pt, fill=yellow!5, inner sep=3pt, font=\scriptsize, text=blue!70!black, below=0.4cm of layer] (mode) {%
  \textbf{Generation:} \texttt{is\_causal=true}, Convection PDE, WavePacketDecoder $\to$ logits%
};
\node[draw, dashed, rounded corners=3pt, fill=green!5, inner sep=3pt, font=\scriptsize, text=green!50!black, below=0.15cm of mode] (mode2) {%
  \textbf{Classification:} \texttt{is\_causal=false}, Diffusion PDE, AttnPool $\to$ Linear $\to$ softmax%
};
\node[draw, dashed, rounded corners=3pt, fill=orange!5, inner sep=3pt, font=\scriptsize, text=orange!60!black, below=0.15cm of mode2] (mode3) {%
  \textbf{Regression:} \texttt{is\_causal=false}, Diffusion PDE, AttnPool $\to$ Linear $\to$ values%
};

\end{tikzpicture}
\caption{Complete LSH-SFA architecture with SpacetimeFieldLayer. The same core layers serve generation (causal mode with convection PDE), classification (bidirectional mode with diffusion PDE + softmax head), and regression (bidirectional mode with diffusion PDE + linear head), differing only in the causal toggle, PDE mode, and output head.}
\label{fig:architecture}
\end{figure*}

\subsection{Unified Decoder-as-Encoder Architecture}

A distinctive property of LSH-SFA is that a single decoder-only architecture serves classification, regression, and generation tasks, without requiring separate encoder and decoder stacks as in BERT-style or BART-style models. This unification is enabled by three architectural mechanisms that switch behavior at the layer level:

\textbf{Bidirectional/Causal Toggle.} Each SpacetimeFieldLayer accepts an \texttt{is\_causal} flag. For classification and regression, all layers operate with \texttt{is\_causal = false}, allowing every token to attend to the full sequence---effectively converting the decoder into an encoder. For generation, layers use \texttt{is\_causal = true}, enforcing the light-cone causal structure. This is fundamentally different from GPT-style models, where causal masking is hard-coded into the architecture and cannot be disabled without retraining.

\textbf{PDE Mode Selection.} The Spacetime Engine naturally provides task-appropriate field evolution: the diffusion equation ($\alpha \nabla^2 F + \beta F^3$) for classification and regression, where the Laplacian performs bidirectional information aggregation; and the convection equation ($-c \cdot \partial F/\partial x + \beta F^3$) for generation, where backward differences enforce causal information flow. The same layer structure accommodates both modes by switching the PDE discretization scheme. Classification and regression share the diffusion PDE because both require global context understanding---the key difference lies only in the output head (softmax vs.\ linear).

\textbf{Task-Adaptive Pooling and Output Heads.} For generation, the WavePacketDecoder projects the final hidden state at each position to vocabulary logits. For classification and regression, we replace the token-level decoder with an attention pooling mechanism followed by a task-specific head:

\begin{equation}
\text{scores} = \text{Linear}_{d \to 1}(\mathbf{H}), \quad \mathbf{w} = \text{softmax}(\text{scores}, \text{dim}=1), \quad \mathbf{z} = \sum_i w_i \mathbf{h}_i
\end{equation}

where $\mathbf{H} \in \mathbb{R}^{n \times d}$ is the sequence of hidden states. The learned query vector assigns importance weights to each position, producing a fixed-dimensional representation $\mathbf{z} \in \mathbb{R}^d$. The task-specific output is then:
\begin{itemize}
\item \textbf{Classification:} $\hat{\mathbf{y}} = \text{softmax}(\text{Linear}_{d \to C}(\mathbf{z}))$, trained with cross-entropy loss
\item \textbf{Regression:} $\hat{\mathbf{y}} = \text{Linear}_{d \to T}(\mathbf{z})$, trained with MSE loss, where $T$ is the number of target variables
\item \textbf{Generation:} $\hat{\mathbf{y}}_i = \text{WavePacketDecoder}(\mathbf{h}_i)$ for each position $i$, trained with next-token cross-entropy loss
\end{itemize}

This attention pooling outperforms simple mean pooling by learning which positions are most informative, and outperforms last-token pooling (used by GPT-style models) because in bidirectional mode no single position has privileged access to the full sequence.

\begin{table}[h]
\centering
\caption{Comparison of unified architecture strategies.}
\label{tab:unified_arch}
\begin{tabular}{lccc}
\toprule
\textbf{Property} & \textbf{BERT/GPT} & \textbf{BART/T5} & \textbf{LSH-SFA} \\
\midrule
Architecture & Encoder or Decoder & Encoder-Decoder & Unified Decoder \\
Classification & Encoder only & Encoder only & Decoder (bidirectional) \\
Regression & Encoder only & Encoder only & Decoder (bidirectional) \\
Generation & Decoder only & Encoder-Decoder & Decoder (causal) \\
Shared weights & No & Partial & Yes (core layers) \\
Causal toggle & Hard-coded & Hard-coded & Per-layer switch \\
Task pooling & [CLS] / Last token & Encoder output & Attention pooling \\
Output head & Task-specific & Task-specific & Task-specific \\
Loss function & CE / MSE & CE / MSE & CE / MSE / CE \\
\bottomrule
\end{tabular}
\end{table}

This unified design offers practical advantages: (1) a single model checkpoint can serve all three task types by switching the forward path, (2) pretraining on language modeling naturally initializes the bidirectional mode for downstream classification and regression, and (3) the physical structure (wave packets, field evolution) remains consistent across tasks, enabling transfer of learned physical priors.

\section{Experiments}

\subsection{AG News Text Classification}

\begin{table}[h]
\centering
\caption{Model configurations for AG News classification.}
\label{tab:agnews_config}
\begin{tabular}{lcc}
\toprule
\textbf{Hyperparameter} & \textbf{LSH-SFA} & \textbf{Transformer} \\
\midrule
$d_{\text{field}}$ / $d_{\text{model}}$ & 256 & 256 \\
Number of layers & 2 & 4 \\
Attention heads & 64 & 8 \\
$d_{\text{time}}$ & 64 & --- \\
Evolution steps & 2 & --- \\
FFN hidden dim & --- & 1024 \\
Total parameters & \textbf{1,024,276} & \textbf{1,194,756} \\
Parameter reduction & \textbf{14\%} & --- \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{AG News classification accuracy.}
\label{tab:agnews_results}
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{Layers} & \textbf{Parameters} & \textbf{Accuracy} \\
\midrule
Transformer & 4 & 1,194,756 & 91.41\% \\
LSH-SFA & 2 & 1,024,276 & \textbf{91.80\%} \\
\bottomrule
\end{tabular}
\end{table}

A 2-layer LSH-SFA matches a 4-layer Transformer, demonstrating 2 additional layers' worth of information density from physical priors. Notably, this result is achieved using the decoder-only architecture in bidirectional mode (Section~8.1): all layers use \texttt{is\_causal = false} with the diffusion PDE, and attention pooling aggregates the sequence into a fixed-dimensional representation for classification. The same model structure, when switched to causal mode with the convection PDE, serves as a language model---demonstrating the unified architecture's effectiveness across task types.

\subsection{WikiText-2 Language Modeling}

\begin{table}[h]
\centering
\caption{WikiText-2 perplexity comparison.}
\label{tab:wikitext_results}
\begin{tabular}{lcc}
\toprule
\textbf{Model} & \textbf{Best PPL} & \textbf{Final PPL} \\
\midrule
Transformer & 1069 & 1069 \\
LSH-SFA & \textbf{654} & \textbf{654} \\
\bottomrule
\end{tabular}
\end{table}

Critical observation: LSH-SFA shows \textbf{no overfitting} (best PPL = final PPL), attributed to energy normalization acting as an implicit regularizer.

\section{Complexity Analysis}

\begin{table}[h]
\centering
\caption{Computational complexity comparison.}
\label{tab:complexity}
\begin{tabular}{lcc}
\toprule
\textbf{Operation} & \textbf{Transformer} & \textbf{LSH-SFA} \\
\midrule
Attention & $O(n^2 d)$ & $O(n d)$ \\
Causal masking & $O(n^2)$ & $O(n \log n)$ \\
FFN / Evolution & $O(n d^2)$ & $O(n d^2 + n d \cdot T)$ \\
Generation per step & $O(n d)$ & $O(d)$ \\
\bottomrule
\end{tabular}
\end{table}

where $n$ is sequence length, $d$ is model dimension, and $T$ is evolution steps (default 2).

\begin{table}[h]
\centering
\caption{Spacetime Engine vs.\ Transformer FFN.}
\label{tab:ffn_comparison}
\begin{tabular}{lcc}
\toprule
\textbf{Property} & \textbf{Transformer FFN} & \textbf{Spacetime Engine} \\
\midrule
Physical interpretation & None & PDE-based \\
Causality & None (same for all) & Mode-dependent \\
Hidden dimension & $4 \times d_{\text{model}}$ & $d_{\text{field}}$ \\
Task adaptation & Shared weights & Diffusion/Convection switch \\
Gradient stability & LayerNorm & Energy normalization \\
\bottomrule
\end{tabular}
\end{table}

\section{Related Work}

\textbf{Linear attention.} Linear Transformer \citep{katharopoulos2020transformers} and Performer \citep{choromanski2021rethinking} reduce attention to $O(n)$ via kernel approximations. LSH-SFA differs in using Gaussian similarity with learnable bandwidth and frequency-adaptive decay.

\textbf{State-space models.} Mamba \citep{gu2023mamba} and S4 \citep{gu2021efficiently} achieve $O(n)$ sequence modeling. LSH-SFA derives the linear recurrence from physical wave propagation rather than learned state transitions.

\textbf{Physics-informed neural networks.} PINNs \citep{raissi2019physics} use PDEs as constraints. LSH-SFA goes further by replacing internal computations with PDE evolution, making the PDE the computation rather than a constraint.

\textbf{Ginzburg-Landau theory.} The nonlinear term $\beta F^3$ in our PDE formulation is inspired by the Ginzburg-Landau equation~\citep{ginzburg1950k}, which models phase transitions in condensed matter physics. In the original framework, the cubic term arises from the free energy expansion near a critical point and provides nonlinear self-interaction that balances the linear diffusion. This connection provides a principled physical basis for the nonlinear term, distinguishing it from arbitrary polynomial choices.

\textbf{Neural ODEs.} Neural ODEs~\citep{chen2018neural} parameterize the derivative of hidden states. LSH-SFA uses physically motivated PDEs (diffusion, convection, wave) rather than learned dynamics, providing interpretability and task-appropriate inductive biases.

\textbf{Gaussian processes.} Gaussian Processes~\citep{rasmussen2006gaussian} provide a probabilistic framework for function approximation using kernel methods. LSH-SFA's Gaussian similarity mechanism can be understood as a neural instantiation of the Gaussian kernel, where wave packets serve as adaptive basis functions~\citep{broomhead1988multivariable}. The ELU+1 kernel mapping follows~\citep{clevert2016fast, tsai2019transformer}.

\textbf{Diffusion models.} Score-based diffusion~\citep{song2021scorebased} uses diffusion processes for generative modeling. LSH-SFA uses diffusion for bidirectional information aggregation in classification, a fundamentally different application.

\section{Conclusion}

We have presented LSH-SFA, a physics-informed neural architecture that replaces each core Transformer component with a physically grounded alternative:

\begin{enumerate}
\item \textbf{Wave Packets} replace QKV with interpretable $(k, w, \omega, \theta)$ representations
\item \textbf{Light-Cone Attention} replaces hard masking with frequency-adaptive exponential decay
\item \textbf{PDE Evolution} replaces FFN with diffusion/convection dynamics inspired by the Ginzburg-Landau equation, with proven stability guarantees
\item \textbf{Resonance Coupling} provides frequency-dependent information transfer
\item \textbf{Unified Decoder-as-Encoder} enables a single decoder-only architecture to serve classification (bidirectional mode + diffusion PDE + attention pooling + softmax), regression (bidirectional mode + diffusion PDE + attention pooling + linear head), and generation (causal mode + convection PDE + wave packet decoding) through per-layer causal toggling and PDE mode selection, eliminating the need for separate encoder and decoder stacks
\end{enumerate}

Experiments demonstrate that physical priors provide significant benefits:
\begin{itemize}
\item 14\% parameter reduction while matching accuracy
\item 39\% perplexity improvement on language modeling
\item No overfitting due to energy normalization
\end{itemize}

Future work includes scaling to billions of parameters, integrating the wave equation for oscillatory dynamics, and deeper integration with the LSH semantic architecture~\cite{lsh-three-layer}.

\section{Validation Roadmap}

The current experiments demonstrate proof-of-concept on small-scale benchmarks. To establish LSH-SFA as a competitive architecture, we plan the following validation steps:

\subsection{Short-Term (3 months)}

\begin{enumerate}
\item \textbf{Fair baseline comparison with linear attention models}: Train Mamba~\citep{gu2023mamba} and S4~\citep{gu2021efficiently} baselines under identical configurations (1M parameters, same training epochs, same optimizer) on AG News and WikiText-2 to isolate the contribution of physical priors vs.\ linear recurrence
\item \textbf{Medium-scale language modeling}: Evaluate on WikiText-103 and Penn Treebank to assess whether the 39\% perplexity improvement scales beyond WikiText-2
\item \textbf{Scaling curve analysis}: Train LSH-SFA at 1M, 4M, 16M, and 64M parameters on the same dataset to characterize the scaling behavior and compare with Transformer scaling laws
\end{enumerate}

\subsection{Medium-Term (6 months)}

\begin{enumerate}
\item \textbf{Billion-parameter pretraining}: Pretrain a 1B-parameter LSH-SFA on a standard corpus (e.g., The Pile or RedPajama) to evaluate training stability and convergence at scale
\item \textbf{Wave equation integration}: Implement and validate the leapfrog discretization scheme for the wave equation, targeting tasks with periodic structure (music, code)
\item \textbf{Multi-modal extension}: Extend wave packet representations to image patches, where $k$ encodes spatial frequency and $\omega$ encodes temporal dynamics for video
\end{enumerate}

\subsection{Long-Term (12 months)}

\begin{enumerate}
\item \textbf{Hardware co-design}: Optimize the parallel prefix scan and PDE evolution for GPU tensor cores and potential quantum annealing backends
\item \textbf{Embodied intelligence validation}: Integrate with the LSH semantic architecture~\cite{lsh-three-layer} for real-time perception-action loops in robotics
\end{enumerate}

\bibliographystyle{plainnat}
\begin{thebibliography}{10}

\bibitem[Broomhead \& Lowe(1988)]{broomhead1988multivariable}
David~S. Broomhead and David Lowe.
\newblock Multivariable functional interpolation and adaptive networks.
\newblock \emph{Complex Systems}, 2(3):321--355, 1988.

\bibitem[Chen et~al.(2018)]{chen2018neural}
Ricky T.~Q.~Chen, Yulia Rubanova, Jesse Bettencourt, and David Duvenaud.
\newblock Neural ordinary differential equations.
\newblock In \emph{NeurIPS}, 2018.

\bibitem[Choromanski et~al.(2021)]{choromanski2021rethinking}
Krzysztof Choromanski, Valerii Likhosherstov, David Dohan, Xingyou Song, Andreea Gane, Tamas Sarlos, Peter Hawkins, Jared Davis, Afroz Mohiuddin, Lukasz Kaiser, et~al.
\newblock Rethinking attention with performers.
\newblock In \emph{ICLR}, 2021.

\bibitem[Clevert et~al.(2016)]{clevert2016fast}
Djork-Arn\'{e} Clevert, Thomas Unterthiner, and Sepp Hochreiter.
\newblock Fast and accurate deep network learning by exponential linear units (ELUs).
\newblock In \emph{ICLR}, 2016.

\bibitem[Ginzburg \& Landau(1950)]{ginzburg1950k}
V.~L. Ginzburg and L.~D. Landau.
\newblock On the theory of superconductivity.
\newblock \emph{Zh. Eksp. Teor. Fiz.}, 20:1064--1082, 1950.

\bibitem[Gu et~al.(2021)]{gu2021efficiently}
Albert Gu, Karan Goel, and Christopher R{\'e}.
\newblock Efficiently modeling long sequences with structured state spaces.
\newblock In \emph{ICLR}, 2021.

\bibitem[Gu \& Dao(2023)]{gu2023mamba}
Albert Gu and Tri Dao.
\newblock Mamba: Linear-time sequence modeling with selective state spaces.
\newblock \emph{arXiv preprint arXiv:2312.00752}, 2023.

\bibitem[Katharopoulos et~al.(2020)]{katharopoulos2020transformers}
Angelos Katharopoulos, Apoorv Vyas, Nikolaos Pappas, and Fran{\c{c}}ois Fleuret.
\newblock Transformers are RNNs: Fast autoregressive transformers with linear attention.
\newblock In \emph{ICML}, 2020.

\bibitem[Li(2026a)]{lsh-three-layer}
Hengbo Li.
\newblock LSH: Element-Observers Semantic Architecture with Emergent Execution.
\newblock \emph{LSH Protocol Preprint}, 2026.

\bibitem[Raissi et~al.(2019)]{raissi2019physics}
Maziar Raissi, Paris Perdikaris, and George~E Karniadakis.
\newblock Physics-informed neural networks: A deep learning framework for solving forward and inverse problems involving nonlinear partial differential equations.
\newblock \emph{Journal of Computational Physics}, 378:686--707, 2019.

\bibitem[Rasmussen \& Williams(2006)]{rasmussen2006gaussian}
Carl~Edward Rasmussen and Christopher K.~I. Williams.
\newblock \textit{Gaussian Processes for Machine Learning}.
\newblock MIT Press, 2006.

\bibitem[Song et~al.(2021)]{song2021scorebased}
Yang Song, Jascha Sohl-Dickstein, Diederik P.~Kingma, Abhishek Kumar, Stefano Ermon, and Ben Poole.
\newblock Score-based generative modeling through stochastic differential equations.
\newblock In \emph{ICLR}, 2021.

\bibitem[Tsai et~al.(2019)]{tsai2019transformer}
Yi-Hsiu Tsai, Shaojie Bai, Paul Pu, and Ruslan Salakhutdinov.
\newblock Transformer dissection: An unified understanding for transformer's attention via the lens of kernel.
\newblock In \emph{EMNLP}, 2019.

\bibitem[Vaswani et~al.(2017)]{vaswani2017attention}
Ashish Vaswani, Noam Shazeer, Niki Parmar, Jakob Uszkoreit, Llion Jones, Aidan~N Gomez, {\L}ukasz Kaiser, and Illia Polosukhin.
\newblock Attention is all you need.
\newblock In \emph{NeurIPS}, 2017.

\end{thebibliography}

\end{document}
